Reproducible methods for modeling combined public transport and cycling trips and associated benefits: evidence from the biclaR tool

### 2024-10-06 paper/PaperCEUS/PaperCEUS_rev2_word.tex
tex minor updates

Using the
\href{https://github.com/dabreegster/odjitter}{\texttt{odjitter} R
package}, we disaggregated the OD data into desire lines reprenting a
maximum of 100 trips each. Figure \ref{fig:jitter} illustrates the
contrast between trip representation through the traditional method,
which connects a single desire line between each district, and the
presentation achieved through the randomization and disaggregation of
trips between districts, specifically for the Lisbon metropolitan area.
As shown, the city of Lisbon (in the centre) is the main attractor of
trips, with a high number of trips to and from the other municipalities.

@@

The mobility survey collects the origin and destination of trips but
does not include the respective routes. Modeling the realistic
cycling-PT routes between OD pairs depends on assumptions regarding the
characteristics of the cycling and road networks and the location of
public transport interfaces. Other constraints regarding the behavior of
potential cyclists determine the routing results. For example, such
restrictions can favor low speed, low traffic streets, more direct
routes, and less steep paths, among others, which are suitable for
cycling.

@@

To estimate the car emissions, we used the EMEP/EEA's COPERT software v5
methods and reference values \citep{COPERT} for a Tier 3 detail level.
We used a family-size vehicle, EURO standard, and gasoline or diesel
fuel. All trips were considered to be made under urban conditions and at
an average speed of 15 km/h during rush hour periods. Since the average
distance traveled per trip influences the overconsumption and emissions
from cold-start engine operation, we estimated energy and emission
factors for different ranges of trips at 500-meter intervals.\\
An equation was then used to calculate emission factors for the two
types of fuel, for each type of pollutant, whose explanatory variables
are driving speed (\(speed\), in km/h) and average trip distance
(\(l_{trip}\), in km/trip). Thus, the emission factors
(\(EF_{fuel,l_{trip},speed}\), in g/km) can be calculated using equation
\ref{eq:fatoremissaoauto}.

@@

\caption{\label{tab:summary1}\label{summary1}Summary of the cycling potencial of the intermodality scenario. Values in 'trips/day'.}
\centering
\begin{tabular}[t]{llr>{\raggedleft\arraybackslash}p{7em}>{\raggedleft\arraybackslash}p{7em}}
\toprule
Target & Routing & Total trips & Baseline Cycling + PT & Potencial Cycling + PT\\
\midrule
4\% & safe & 538 514 & 2 312 & 20 385\\
4\% & direct & 500 880 & 2 274 & 18 944\\
10\% & safe & 538 514 & 2 312 & 52 323\\
10\% & direct & 500 880 & 2 274 & 48 609\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summary1b}\label{summary1b}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}}
\toprule
Target & Routing & Avoided Mortality (deaths/yr) & Social benefits (k€/yr) & Avoided CO2eq (ton/yr) & Environmental benefits (k€/yr)\\
\midrule
4\% & safe & 4.1 & 12 717 & 2 958 & 238\\
4\% & direct & 4.0 & 12 441 & 3 004 & 241\\
10\% & safe & 10.0 & 32 820 & 7 590 & 610\\
10\% & direct & 10.0 & 31 800 & 7 694 & 618\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summary22}\label{summary22}Summary of the avoided emmissions (ton/year) and corresponding monetization (thousand €) by replacing car trips with PT, in the second leg.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}r}
\toprule
Target & Routing & CO2eq & CO & PM10 & NOx & VOC & Value (k€)\\
\midrule
4\% & safe & 8 593 & 17 & 1.9 & 27 & 0.8 & 1 425\\
4\% & direct & 8 702 & 18 & 2.0 & 28 & 0.8 & 1 453\\
10\% & safe & 20 627 & 42 & 4.6 & 65 & 2.0 & 3 431\\
10\% & direct & 20 793 & 42 & 4.7 & 66 & 1.9 & 3 487\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 11 551 & 14 380 & 127 534\\
4\% & direct & 11 706 & 14 135 & 125 016\\
10\% & safe & 28 217 & 36 861 & 325 814\\
10\% & direct & 28 487 & 35 905 & 318 062\\
\bottomrule
\end{tabular}
\end{table}

@@

The environmental impacts represent less than 2\% of the
socio-environmental benefits (in value) from replacing car trips to
bicycle in first-and-last legs. For the PT segment, we did not estimate
the social impacts from substituting car trips. One of the main
socio-environmetal benefits, valued after monetization, comes from the
increase in physical activity \citep{Felix2023ES}. Although there are
also social benefits form shifting car trips to PT, its health benefits
would not be as high as shifting to cycling. The literature shows that
the Metabolic Equivalent Tasks (MET) for ``riding in a bus or a train''
is 1.3 plus the ``walking for transportation'' as 3.5, while ``driving a
car'' is 2.5 \citep{MET2011}. The difference between these activities -
shifting from car to PT - is not very obvious when compared to shifting
from car to cycling, whose MET is about 6.8. Nevertheless, future works
should also encompass the estimation of the social impacts for the PT
leg of the journey, shifting from car.

@@

The findings indicate that cycling combined with public transport has
the potential to account for 10\% of total trips. Transitioning to
cycling for the first and last segments could lead to a decrease in
annual CO\textsubscript{2}eq emissions by a substantial margin, ranging
from 3,000 to 7,500 tons per day, depending on the cycling target and
routing profile. Moreover, for the second trip segment, the shift from
car to public transport contributes to avoiding up to 20,500 tons of
CO\textsubscript{2}eq emissions annually. These changes are estimated to
yield socio-environmental benefits totaling €125 million to €325 million
over a decade. The quantification of such benefits can support
policy-makers in prioritizing interventions to reduce the reliance on
private motorized modes of transportation.

>>>>

Using the
\href{https://github.com/dabreegster/odjitter}{\texttt{odjitter} R
package}, we disaggregated the OD data into desire lines representing a
maximum of 100 trips each. Figure \ref{fig:jitter} illustrates the
contrast between trip representation through the traditional method,
which connects a single desire line between each district, and the
presentation achieved through the randomization and disaggregation of
trips between districts, specifically for the Lisbon metropolitan area.
As shown, the city of Lisbon (in the centre) is the main attractor of
trips, with a high number of trips to and from the other municipalities.

@@

The mobility survey collects the origin and destination of trips but
does not include the respective routes. Modeling the realistic
cycling-PT routes between OD pairs depends on assumptions regarding the
characteristics of the cycling and road networks and the location of
public transport interfaces. Other constraints regarding the behaviour
of potential cyclists determine the routing results. For example, such
restrictions can favour low speed, low traffic streets, more direct
routes, and less steep paths, among others, which are suitable for
cycling.

@@

To estimate the car emissions, we used the EMEP/EEA's COPERT software v5
methods and reference values \citep{COPERT} for a Tier 3 detail level.
We used a family-size vehicle, EURO standard, and gasoline or diesel
fuel. All trips were considered to be made under urban conditions and at
an average speed of 15 km/h during rush hour periods. Since the average
distance travelled per trip influences the overconsumption and emissions
from cold-start engine operation, we estimated energy and emission
factors for different ranges of trips at 500-meter intervals.\\
An equation was then used to calculate emission factors for the two
types of fuel, for each type of pollutant, whose explanatory variables
are driving speed (\(speed\), in km/h) and average trip distance
(\(l_{trip}\), in km/trip). Thus, the emission factors
(\(EF_{fuel,l_{trip},speed}\), in g/km) can be calculated using equation
\ref{eq:fatoremissaoauto}.

@@

\caption{\label{tab:summary1}\label{summary1}Summary of the cycling potential of the intermodality scenario. Values in 'trips/day'.}
\centering
\begin{tabular}[t]{llr>{\raggedleft\arraybackslash}p{7em}>{\raggedleft\arraybackslash}p{7em}}
\toprule
Target & Routing & Total trips & Baseline Cycling + PT & Potential Cycling + PT\\
\midrule
4\% & safe & 538 514 & 2 312 & 20 385\\
4\% & direct & 500 880 & 2 274 & 18 944\\
10\% & safe & 538 514 & 2 312 & 52 323\\
10\% & direct & 500 880 & 2 274 & 48 609\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summary1b}\label{summary1b}Summary of the cycling potential of intermodality scenario and its socio-environmental benefits for the cycling legs.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}}
\toprule
Target & Routing & Avoided Mortality (deaths/yr) & Social benefits (k€/yr) & Avoided CO2eq (ton/yr) & Environmental benefits (k€/yr)\\
\midrule
4\% & safe & 4.1 & 12 717 & 2 958 & 238\\
4\% & direct & 4.0 & 12 441 & 3 004 & 241\\
10\% & safe & 10.0 & 32 820 & 7 590 & 610\\
10\% & direct & 10.0 & 31 800 & 7 694 & 618\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summary22}\label{summary22}Summary of the avoided emissions (ton/year) and corresponding monetization (thousand €) by replacing car trips with PT, in the second leg.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}>{\raggedleft\arraybackslash}p{3.5em}r}
\toprule
Target & Routing & CO2eq & CO & PM10 & NOx & VOC & Value (k€)\\
\midrule
4\% & safe & 8 593 & 17 & 1.9 & 27 & 0.8 & 1 425\\
4\% & direct & 8 702 & 18 & 2.0 & 28 & 0.8 & 1 453\\
10\% & safe & 20 627 & 42 & 4.6 & 65 & 2.0 & 3 431\\
10\% & direct & 20 793 & 42 & 4.7 & 66 & 1.9 & 3 487\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 11 551 & 14 380 & 127 534\\
4\% & direct & 11 706 & 14 135 & 125 016\\
10\% & safe & 28 217 & 36 861 & 325 814\\
10\% & direct & 28 487 & 35 905 & 318 062\\
\bottomrule
\end{tabular}
\end{table}

@@

The environmental impacts represent less than 2\% of the
socio-environmental benefits (in value) from replacing car trips to
bicycle in first-and-last legs. For the PT segment, we did not estimate
the social impacts from substituting car trips. One of the main
socio-environmental benefits, valued after monetization, comes from the
increase in physical activity \citep{Felix2023ES}. Although there are
also social benefits form shifting car trips to PT, its health benefits
would not be as high as shifting to cycling. The literature shows that
the Metabolic Equivalent Tasks (MET) for ``riding in a bus or a train''
is 1.3 plus the ``walking for transportation'' as 3.5, while ``driving a
car'' is 2.5 \citep{MET2011}. The difference between these activities -
shifting from car to PT - is not very obvious when compared to shifting
from car to cycling, whose MET is about 6.8. Nevertheless, future works
should also encompass the estimation of the social impacts for the PT
leg of the journey, shifting from car.

@@

The findings indicate that cycling combined with public transport has
the potential to account for 10\% of total trips. Transitioning to
cycling for the first and last segments could lead to a decrease in
annual CO\textsubscript{2}eq emissions by a substantial margin, ranging
from 3,000 to 7,500 tons per day, depending on the cycling target and
routing profile. Moreover, for the second trip segment, the shift from
car to public transport contributes to avoiding up to 20,500 tons of
CO\textsubscript{2}eq emissions annually. These changes are estimated to
yield socio-environmental benefits totalling €125 million to €325
million over a decade. The quantification of such benefits can support
policy-makers in prioritizing interventions to reduce the reliance on
private motorized modes of transportation.
